\vspace{-0.5cm}The X-shooter GTO sample of GRB afterglow and host Galaxy spectra

### 2016-11-09 document/XSGRB_sample.tex
Skeleton

	\author{
		J.~Selsing\altaffilmark{1}, 
		D.~Malesani\altaffilmark{1}, 
		P.~Goldoni\altaffilmark{14}, 
		T.~Kr\"{u}hler\altaffilmark{1}, 
		J.~P.~U. Fynbo\altaffilmark{1}, 
		A.~de~Ugarte~Postigo\altaffilmark{11}, 
		J.~Japelj\altaffilmark{20},
		P.~D'Avanzo,
		Z.~Cano,
		S.~Covino\altaffilmark{10}, 
		V.~D'Elia\altaffilmark{7, 12}, 
		H.~Flores,
		O.~E.~Hartoog\altaffilmark{6},
		J.~Hjorth\altaffilmark{1}, 
		P.~Jakobsson\altaffilmark{5}, 
		A.~Levan,
		A.~Melandri,
		S.~Piranomonte \altaffilmark{7},
		R.~S\'anchez-Ram\'irez\altaffilmark{11},
		S.~Schulze\altaffilmark{17, 18}, 
		N.~R.~Tanvir\altaffilmark{19},
		C.~Th{\"o}ne,
		S.~D.~Vergani\altaffilmark{7, 8},
		P.~M.~Vreeswijk\altaffilmark{3}, 
		D.~J.~Watson\altaffilmark{1},
		K.~Wiersema\altaffilmark{19},
		D.~Xu\altaffilmark{1}
		L.~Christensen\altaffilmark{1},
		A.~De~Cia\altaffilmark{3}, 
		L.~Kaper\altaffilmark{6}, 
		L.~A.~Antonelli,
		F.~Fiore,
		A.~Gomboc,
		P.~Groot,
		F.~Hammer,
		C.~Ledoux\altaffilmark{2}, 
		E.~Maiorano,
		B.~Milvang-Jensen\altaffilmark{1}, 
		E.~Palazzi,
		E.~Pian,
		J.~Schaye,
		G.~Tagliaferri\altaffilmark{7},
		R.~A.~M.~J.~Wijers\altaffilmark{6}
	}

@@

@@

	\section{Results}
	\input{tables/burst_overview.tex}	
	\subsection{Spectral resolution}
	The afterglow spectra described in this paper are obtained in
	Target-of-Opportunity (override) mode.
	In most cases there is therefore little possibility to tweak slit widths to the
	seeing at the time of observations (i.e. to optimise spectral resolution and
	signal to noise), and almost all our data is therefore taken with a fixed set
	of slit widths and binning, described above.
	In a fair number of cases, the seeing full width at half maximum (FWHM) is
	considerably smaller than the silt width, and the delivered spectral resolution
	will then be determined by the seeing rather than slit width, as afterglows are
	point sources (this is evidently not the case for extended sources, e.g for
	host galaxies).
	The delivered resolution for slit width dominated spectra post-reduction and
	extraction can easily be determined from the bright sky emission lines.
	For afterglow spectra with very high signal to noise, the delivered spectral
	resolution can at times be determined from the science data themselves.
	However, in the presence of multiple velocity components in absorption, other
	forms of line broadening, and a lack of lines at some redshifts, this is
	difficult to do at poorer signal to noise ratios (the majority of spectra in
	our sample).
	A broad starting value for the expect resolution will help fitting of these
	spectra, and can be important in upper limit determination, and for this reason
	we construct a aim to construct a crude relation between the seeing and the
	delivered resolution at our slit width, binning, and reduction pipeline
	settings.
	To this end we use observations of telluric standard stars that are taken with
	identical instrument settings as our afterglow spectra, usually just after the
	science data, as part of the ESO X-shooter calibration plan.
	These spectra have been reduced together with the afterglow spectra, using
	identical pipeline settings with the same version of the pipeline.
	First we fit a Gaussian function in the spatial direction of the trace of the
	standard star at 792 nm (i.e. in the VIS arm).
	After this, we fit a series of  20 telluric absorption lines in the telluric
	standard star spectra with Gaussians, taking care to select transitions that
	are not almost-resolved multiples, should be intrinsically unresolved, and are
	in areas with well defined continuum flux.
	We pick 34 telluric standard stars spanning a range of DIMM seeing values, with
	the majority between $0.5-1.5 arcsec$.
	The resulting distribution of spectral FWHM (km/s) as a function of spatial
	FWHM at 792 nm is fairly well described by a linear relation $a + b*x$, with
	$x$ the spatial FWHM in pixels (with 0.15 arc sec per pixel),  $a= 21.4 \pm
	1.3$ km/s, $b=1.4 \pm 0.2$.
	We use this linear relation as a way to estimate the spectral resolution for
	medium to poor signal to noise afterglow spectra in the VIS arm.
	To extend this to the UVB and NIR arm, we measured a series of lines in NIR arm
	spectra of a subset of 19 sources used for the VIS arm above, and find that the
	resulting distribution is consistent with a simple scaling of the VIS arm
	relation by the ratio of resolutions of the NIR and VIS arm for unresolved,
	slit filling, sources as given on the ESO instrument website.
	The UVB arm contains no suitable absorption lines to use, and we therefore use
	a scaled value as in the NIR arm.
	While this simple method is not terribly accurate (for one, the spatial profile
	of the trace is not a perfect Gaussian), but it gives a sufficiently accurate
	estimate for the analysis of these poor signal to noise science spectra.

@@

	\subsection{Redshifts}
%	\subsection{Notes on Individual objects}

@@

>>>>

	\author{
		J.~Selsing\altaffilmark{1}, 
		D.~Malesani\altaffilmark{1}, 
		T.~Kr\"{u}hler\altaffilmark{1}, 
		P.~Goldoni\altaffilmark{14}, 
		J.~P.~U. Fynbo\altaffilmark{1}, 
		A.~de~Ugarte~Postigo\altaffilmark{11}, 
		J.~Japelj\altaffilmark{20},
		P.~D'Avanzo,
		Z.~Cano,
		S.~Covino\altaffilmark{10}, 
		V.~D'Elia\altaffilmark{7, 12}, 
		H.~Flores,
		O.~E.~Hartoog\altaffilmark{6},
		J.~Hjorth\altaffilmark{1}, 
		P.~Jakobsson\altaffilmark{5}, 
		A.~Levan,
		A.~Melandri,
		S.~Piranomonte \altaffilmark{7},
		R.~S\'anchez-Ram\'irez\altaffilmark{11},
		S.~Schulze\altaffilmark{17, 18}, 
		N.~R.~Tanvir\altaffilmark{19},
		C.~Th{\"o}ne,
		S.~D.~Vergani\altaffilmark{7, 8},
		P.~M.~Vreeswijk\altaffilmark{3}, 
		D.~J.~Watson\altaffilmark{1},
		K.~Wiersema\altaffilmark{19},
		D.~Xu\altaffilmark{1}
		L.~Christensen\altaffilmark{1},
		A.~De~Cia\altaffilmark{3}, 
		L.~Kaper\altaffilmark{6}, 
		L.~A.~Antonelli,
		F.~Fiore,
		A.~Gomboc,
		P.~Groot,
		F.~Hammer,
		C.~Ledoux\altaffilmark{2}, 
		E.~Maiorano,
		B.~Milvang-Jensen\altaffilmark{1}, 
		E.~Palazzi,
		E.~Pian,
		J.~Schaye,
		G.~Tagliaferri\altaffilmark{7},
		R.~A.~M.~J.~Wijers\altaffilmark{6}
	}

@@

	We simultaneously want to minimize any biases against astrophysical conditions while at the same time maximizing likelihood of observations. By restricting the selection criteria to conditions local to Galaxy.

@@

	\subsection{Spectral resolution}
	The afterglow spectra described in this paper are obtained in
	Target-of-Opportunity (override) mode.
	In most cases there is therefore little possibility to tweak slit widths to the
	seeing at the time of observations (i.e. to optimise spectral resolution and
	signal to noise), and almost all our data is therefore taken with a fixed set
	of slit widths and binning, described above.
	In a fair number of cases, the seeing full width at half maximum (FWHM) is
	considerably smaller than the silt width, and the delivered spectral resolution
	will then be determined by the seeing rather than slit width, as afterglows are
	point sources (this is evidently not the case for extended sources, e.g for
	host galaxies).
	The delivered resolution for slit width dominated spectra post-reduction and
	extraction can easily be determined from the bright sky emission lines.
	For afterglow spectra with very high signal to noise, the delivered spectral
	resolution can at times be determined from the science data themselves.
	However, in the presence of multiple velocity components in absorption, other
	forms of line broadening, and a lack of lines at some redshifts, this is
	difficult to do at poorer signal to noise ratios (the majority of spectra in
	our sample).
	A broad starting value for the expect resolution will help fitting of these
	spectra, and can be important in upper limit determination, and for this reason
	we construct a aim to construct a crude relation between the seeing and the
	delivered resolution at our slit width, binning, and reduction pipeline
	settings.
	To this end we use observations of telluric standard stars that are taken with
	identical instrument settings as our afterglow spectra, usually just after the
	science data, as part of the ESO X-shooter calibration plan.
	These spectra have been reduced together with the afterglow spectra, using
	identical pipeline settings with the same version of the pipeline.
	First we fit a Gaussian function in the spatial direction of the trace of the
	standard star at 792 nm (i.e. in the VIS arm).
	After this, we fit a series of  20 telluric absorption lines in the telluric
	standard star spectra with Gaussians, taking care to select transitions that
	are not almost-resolved multiples, should be intrinsically unresolved, and are
	in areas with well defined continuum flux.
	We pick 34 telluric standard stars spanning a range of DIMM seeing values, with
	the majority between $0.5-1.5 arcsec$.
	The resulting distribution of spectral FWHM (km/s) as a function of spatial
	FWHM at 792 nm is fairly well described by a linear relation $a + b*x$, with
	$x$ the spatial FWHM in pixels (with 0.15 arc sec per pixel),  $a= 21.4 \pm
	1.3$ km/s, $b=1.4 \pm 0.2$.
	We use this linear relation as a way to estimate the spectral resolution for
	medium to poor signal to noise afterglow spectra in the VIS arm.
	To extend this to the UVB and NIR arm, we measured a series of lines in NIR arm
	spectra of a subset of 19 sources used for the VIS arm above, and find that the
	resulting distribution is consistent with a simple scaling of the VIS arm
	relation by the ratio of resolutions of the NIR and VIS arm for unresolved,
	slit filling, sources as given on the ESO instrument website.
	The UVB arm contains no suitable absorption lines to use, and we therefore use
	a scaled value as in the NIR arm.
	While this simple method is not terribly accurate (for one, the spatial profile
	of the trace is not a perfect Gaussian), but it gives a sufficiently accurate
	estimate for the analysis of these poor signal to noise science spectra.

@@

	\section{Results}
	\input{tables/burst_overview.tex}

	\subsection{Complete sample}

	\subsection{Follow-up timing and afterglow brightness}
	This sections should contain a plot og the follow-up delay and the acquisition image brightness at the time of observations.

	\subsection{Properties of rejected triggers}
	How many bursts have been excluded due to particularities in the follow-up attempt. Visitor, Weather, technical issues.

	\subsection{On the nature of triggered GRBs without detected optical afterglow.}
	This should include observations where the acquisition image has been taken, but observations were discontinued due to non-detection. What are the X-ray properties of these bursts?

	\subsection{Redshifts}

	\subsubsection{Afterglow-based redshifts}

	\subsubsection{Emissionline-based redshifts}	
	Redshift completeness

	\subsection{Hydrogen column densities}

	\subsubsection{Comparison to X-ray determined columns}

@@

	\subsection{Comparison to previous samples.}	
	This should also include a discussion of dark bursts. 
	Cuchiara. BAT6 (Salvaterra), TOUGH

	\subsection{Comparison to previous samples.}

	\subsection{X-ray column densities vs. optically derived ones}

	\subsection{Why is the HI column truncated?}

	What is the expected transmission for a high-column HI, where the hydrogen primarily will be in the form of H2?

	\section{Conclusion}
